D.I.A.G.R.A.M.: Development of Image Analysis for Graph Recognition And Modeling

### 2025-06-29 doc/article/article.tex
Update paper
Edited paper
Edit paper

@@

\subsection{Extractor Module}
\subsubsection{Preprocessing}
TODO\\
\subsubsection{Bounding Box Detection}
TODO\\
\subsubsection{Content Recognition}
After having determined all the bounding boxes of the diagram image, in order to correctly create the diagram internal representation, there's the need to link together text, elements and relations.\\

\subsubsection{Arrow Orientation}
To later correctly assign arrows to their source and target elements, the need to determine where the tip of an arrow lies. This task involves a few problems, among which determining exactly the arrow shape and its end - because sometimes other pen signs may be found in the bounding box of the arrow itself. [TODO: Dire l'approccio finale usato]. \\

@@

\subsubsection{Relation-Text Association}
After a text box is assigned to a relation, some computation is needed to fully understand how the text is associated to the relation itself. [TODO: Come trova source e target]. The relation bounding box is then split into three different sections. The overlap between the text bounding box and each of the three sections boxes is computed - then, the section with the highest overlap value is then assigned the text under consideration.

@@

\subsection{Arrow Orientation}
As we've seen in the previous section, the need to find the exact position of an arrow's head arises (see \textbf{Appendix~\ref{double_clustering}} for a full overview of the first approach).

TODO: Current approach.

@@

\bibliographystyle{IEEEtran}
\bibliography{references}

@@

However, this approach didn't work as expected (see Figure~\ref{fig:template_matching_fail}). [TODO: Spiegare perchè].

@@

>>>>

\section{Previous Works}
\cite{arrowrcnn} \\
\cite{interactiveUML} \\
\cite{sketchdiagram}

@@

\subsection{Extractor Module}

\begin{figure}[H]
	\centering
	\includegraphics[width=\linewidth]{flowchart-extractor.png}
	\caption{The complete pipeline of the extractor module.}
	\label{fig:flowchart-extractor}
\end{figure}

\subsubsection{Preprocessing}
After some brief experiments, we found out that the object detection task through bounding boxes worked very well without any preprocessing technique. Moreover, the use of certain filters actually \textit{downgraded} the performances - e.g. the use of aggressive median filters reduced the capacity of the OCR model for text extraction.

The only preprocessing applied is a gray-scale transformation. Other transformations are not deemed necessary - the model employed automatically resize the input image, so it's not needed to do it beforehand. \\

\subsubsection{Bounding Box Detection}
\label{extr:bbd}
TODO: FasterRCNN - backbone ResNet 50FPN~\cite{rcnn}.  \\ After having determined all the bounding boxes of the diagram image, in order to correctly create the diagram internal representation, there's the need to link together text, elements and relations.\\

\subsubsection{Arrow Orientation}
To later correctly assign arrows to their source and target elements, the need to determine where the tip of an arrow lies. This task involves a few problems, among which determining exactly the arrow shape and its end - because sometimes other pen signs may be found in the bounding box of the arrow itself. After a few trials (and errors; for further insight, refer to section~\ref{exp:arrow_orientation}), the approach yielding the best results was through the use of the object detection network (see section~\ref{extr:bbd}) to recognize both arrow heads and tails. \\

@@

\subsubsection{Relation-Text Association}
After a text box is assigned to a relation, some computation is needed to fully understand how the text is associated to the relation itself. The relation bounding box is then split into three different sections. The overlap between the text bounding box and each of the three sections boxes is computed - then, the section with the highest overlap value is then assigned the text under consideration. \\

\subsubsection{Relationships Building}
Once the text has been dealt with, the last step is to effectively compute which elements are connected by which arrow. This is done by finding, for each arrow, the elements nearer to the head (which we will assume to be the 'target' of the relation) and to the tail (the 'source' of the relationship exemplified by the arrow). These associations are nullified if the distance is above a certain adjustable threshold. This last step is done under the hypothesis that an arrow may come from nothing - or may point to nothing, having actually only either the source or the target.

\section{Dataset}
TODO: Elencare i dataset usati, in che modo erano usati originariamente e come lo usiamo noi (quali dati per quali componenti) \\

@@

\subsection{Arrow Orientation}
\label{exp:arrow_orientation}
As we've seen in the previous section, the need to find the exact position of an arrow's head arises (see \textbf{Appendix~\ref{double_clustering}} for a full overview of the first approaches).

The solution, as stated before, was to

@@

\bibliographystyle{IEEEtran}
\bibliography{article}

@@

However, this approach didn't work as expected (see Figure~\ref{fig:template_matching_fail}). In particular, it never worked around self arrows: the maximized distance inside one of the two clusters is never between the tail and the head of the arrow.

@@

\section{Arrow-Net}
\label{arrow-net}
